---
title: "Basic Utility Functions"
author: "Nika Karimi"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
---

# Introduction

The **regexcite** package provides simple yet handy utility functions for common tasks. In this vignette, we demonstrate two core functions:

1. `is_leap(year)`: Determine whether a given integer year is a leap year.
2. `is_palindrome(word)`: Check if a string reads the same forwards and backwards, ignoring spaces, punctuation, and letter case.

Throughout this document, we load the package, call each function with representative examples, and discuss their behavior.



# Loading the Package

Ensure you have installed **regexcite** from CRAN or GitHub:

```r
# Install from CRAN
install.packages("regexcite")

# Or install the development version
# remotes::install_github("<your-username>/regexcite")
```

Load the library:

```{r load-package, message=FALSE}
library(regexcite)
```


# Checking for Leap Years

The `is_leap()` function implements the Gregorian calendar rules:

- Years divisible by 4 → leap years.
- Years divisible by 100 → **not** leap years, unless also divisible by 400.

## Usage

```r
?is_leap
```

## Examples

```{r leap-examples}
years <- c(1992, 1900, 2000, 2023)
leap_flags <- is_leap(years)
data.frame(
  year = years,
  is_leap = leap_flags
)
```

Expected output:

| year | is_leap |
|------|---------|
| 1992 | TRUE    |
| 1900 | FALSE   |
| 2000 | TRUE    |
| 2023 | FALSE   |


# Detecting Palindromes

The `is_palindrome()` function tests whether a single-character string is a palindrome. It:

1. Converts to lowercase.
2. Removes all non-alphanumeric characters.
3. Compares the cleaned string to its reverse.

## Usage

```r
?is_palindrome
```

## Examples

```{r palindrome-examples}
words <- c(
  "radar",
  "A man, a plan, a canal: Panama",
  "Hello, world"
)

pal_flags <- vapply(words, is_palindrome, logical(1))
data.frame(
  word = words,
  is_palindrome = pal_flags
)
```

Expected output:

| word                               | is_palindrome |
|------------------------------------|---------------|
| radar                              | TRUE          |
| A man, a plan, a canal: Panama    | TRUE          |
| Hello, world                       | FALSE         |


# Function Internals

For reference, here are the core implementations:

```r
is_leap <- function(year) {
  if (!is.numeric(year) || any(year %% 1 != 0)) {
    stop("Input must be an integer.")
  }
  (year %% 4 == 0 & year %% 100 != 0) | (year %% 400 == 0)
}

is_palindrome <- function(word) {
  if (!is.character(word) || length(word) != 1) {
    stop("Input must be a single character string.")
  }
  cleaned <- stringr::str_to_lower(
    stringr::str_remove_all(word, "[^[:alnum:]]")
  )
  cleaned == paste(rev(strsplit(cleaned, "")[[1]]), collapse = "")
}
```


# Conclusion

The **regexcite** package’s utility functions simplify common string and date checks. This vignette illustrated how to:

- Identify leap years with `is_leap()`.
- Detect palindromes with `is_palindrome()`.

We encourage you to explore these functions further and contribute enhancements on GitHub. Happy coding!
